% A skeleton file for producing Computer Engineering reports
% https://kgcoe-git.rit.edu/jgm6496/KGCOEReport_template

\documentclass[CMPE]{KGCOEReport}

% The following should be changed to represent your personal information
\newcommand{\classCode}{CMPE 110}  % 4 char code with number
\newcommand{\name}{Andrei Tumbar}
\newcommand{\LabSectionNum}{05}
\newcommand{\LabInstructor}{Mark Byers}	% The slash is to tell LaTeX that the period is between words
												% not sentences so it spaces correctly. It won't appear in the
												% final pdf
\newcommand{\TAs}{Rickie Yeung \\ Tommy Li}
\newcommand{\LectureSectionNum}{01}
\newcommand{\LectureInstructor}{Dr.\ Soldavini}
\newcommand{\exerciseNumber}{2}
\newcommand{\exerciseDescription}{Ohm's Law}
\newcommand{\dateDone}{September 05, 2019}
\newcommand{\dateSubmitted}{September 19, 2019}

\usepackage{circuitikz}
\usepackage{multirow}
\usepackage{float}
\usepackage{array}
\usepackage{pgfplots, pgfplotstable}

\begin{document}
\maketitle

\section*{Abstract}

Understanding Ohm's Law as well as Kirchoff's loop laws is an important This labratory exercise exposes the relationship between voltage, current, and resistance in both a parallel and series configuration. A powersupply supplied a variable voltage across 3 resistors, two in parrallel and the third in series. An effective resistance was calculated and compared to a measured resistance. Potential drop across resistors was measured at every integer voltage from 2-5V. A Voltage vs. Current graph was generated illustrating the linear relationship between the two variables for ohmic resistors.

\section*{Design Methodology}

The three smallest resistors were determined. Resistors were named $R_1$, $R_2$, and $R_3$ in order by increasing resistance. The actual resistance was measured using the multimeter's ohmeter function (See table 1).


\begin{figure}[h!]
\begin{center}
\begin{circuitikz}
\draw
(0,0) to[battery1, l=V] (0,4) --
(0,4) to[R, l_=$R_1$] (4,4) --
(6,4) to[R, l_=$R_3$] (6,0) -- (4,0)
(4,4) to[R, l_=$R_2$] (4,0) --
(0,0)
;
\end{circuitikz}
\end{center}
\caption{Circuit configuration given $R_1$, $R_2$, and $R_3$ have accending resistances.}
\end{figure}

A circuit was wired on a breadboard such that $R_2$ and $R_3$ were placed in parallel to each other and in series to $R_1$. Figure 1 shows an example circuit in the correct configuration. With the power supply voltage set to 5V, potential drop across $R_1$ and $R_2$ was taken. This step was repeated for 2V, 3V, and 4V across the power supply (table 2).

Current through all three resistors was measured at a 5V total potential (see table 3). An effective resistance for the entire circuit was calculated. First the two parallel branches were combined through the follwoing formula:
\[
R_{eq} = \frac{1}{ \sum\frac{1}{R_i} }
\]
Resistance of $R_1$ was added to the value about yielding a total effective resistance of the circuit.  

Finally, current through $R_3$ was measured from 2-5V at 1V intervals from the power supply and recorded (See table 4). Using this data, a graph was generated for Voltage vs. Current through $R_3$ (See figure 2).

\section*{Results and Analysis}
\subsection*{Resistance}
Measured resistance in $R_1$ $R_2$ and $R_3$ showed great accuracy to the rated resistance. Table 1 shows that all measured resistances are with the tolerance rated for each resistor.


\begin{table}[h]
\caption{Measurements taken for all three resistors.}
\begin{center}
\begin{tabular}{cc|c|c}
& \multicolumn{2}{c}{Banded Values} & Measured Values  \\\cline{2-4}
\multirow{2}{*}{}
& Rated Resistance & Rated Tolerance & Actual Resistance \\
& ($\Omega$) & (\%) & ($\Omega$) \\\cline{2-4}
$R_1$ & 330 & $\pm$ 5 & 328.00 \\\cline{2-4}
$R_2$ & 3330 & $\pm$ 5 & 3396.00 \\\cline{2-4}
$R_3$ & 5600 & $\pm$ 5 & 5598.80 \\\cline{2-4}
\end{tabular}
\end{center}
\end{table}

\subsection*{Voltage}
As expected, the voltage across $R_1$ and $R_2$ scale proportionally to the voltage across the power supply. 
{
\renewcommand{\arraystretch}{1.4}
\begin{table}[h]
\caption{Voltage measured across $R_1$ and $R_2$.}
\begin{center}
\begin{tabular}{c>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}}
& $V_{PS}$ & $V_{R_1}$ & $V_{R_2}$ \\\cline{2-4}
\textasciitilde2V & 1.997 & 0.268 & 1.728 \\\cline{2-4}
\textasciitilde3V & 2.997 & 0.402 & 2.592 \\\cline{2-4}
\textasciitilde4V & 3.998 & 0.536 & 3.457 \\\cline{2-4}
\textasciitilde5V & 4.999 & 0.670 & 4.320
\end{tabular}
\end{center}
\end{table}
}

\subsection*{Current}
{
\renewcommand{\arraystretch}{1.4}
\begin{table}[h]
\caption{Voltage measured across $R_1$ and $R_2$.}
\begin{center}
\begin{tabular}{>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}}
$I_{R_1}$ (mA) & $I_{R_2}$ (mA) & $I_{R_3}$ (mA) \\\hline
2.00 & 1.27 & 0.77
\end{tabular}
\end{center}
\end{table}
}

Current through any of the three resistors yielded very low numbers because the effective resistance was so high. Effective resistance came out to be $2441.84 \Omega$ calculated from:

\[
R_{eq} = R_1 + \frac{1}{ \frac{1}{R_2} + \frac{1}{R_3} } = 2441.84\Omega
\]

\subsection*{Results}

Table 4 was generated by measuring the current through $R_3$ and the power supply at every 1V interval from 2-5V. The last column was generated by applying Ohm's Law $V=IR$.

\begin{table}[htb]
\caption{Table to test captions and labels}
\begin{center}
\begin{tabular}{c|>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}}
Voltage & V\textsubscript{PS} & V\textsubscript{R\textsubscript{3}} & I\textsubscript{PS} & I\textsubscript{R\textsubscript{3}} & R\textsubscript{3} \\
(V) & (V) & (V) & (mA) & (mA) & ($\Omega$) \\
\hline
2 & 1.997 & 1.728 & 0.815 & 0.308 & 5616.88  \\
3 & 2.997 & 2.592 & 1.220 & 0.463 & 5593.95 \\
4 & 3.998 & 3.475 & 1.630 & 0.617 & 5607.78 \\
5 & 4.999 & 4.320 & 2.040 & 0.771 & 5603.11
\end{tabular}
\end{center}
\end{table}

\pgfplotstableread{
X Y
0.308 1.728
0.462 2.592
0.617 3.475
0.771 4.320
}\datatable

\begin{figure}[H]
\caption{Voltage over $R_3$ vs Current through $R_3$}
\begin{center}
\begin{tikzpicture}
\begin{axis}[
    xlabel={Current [mA]},
    ylabel={Voltage [V]},
    xmin=0.2, xmax=0.8,
    ymin=1, ymax=5,
    xtick={.2,.4,.6,.8},
    ytick={2,3,4,5},
    legend pos=outer north east,
    ymajorgrids=true,
    grid style=dashed,
]
 
\addplot[only marks, mark = *] table {\datatable};
\addplot [thick] table[
    y={create col/linear regression={y=Y}}
] {\datatable};
\addlegendentry{$data$}
\addlegendentry{%
$\pgfmathprintnumber[precision=3]{\pgfplotstableregressiona} \cdot x
\pgfmathprintnumber[print sign]{\pgfplotstableregressionb}$}

\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}

Figure 2 displays the data from table 4 for $I_{R_3}$ on the X-axis and $V_{R_3}$ on the Y-axis. The slope in figure 2 represents the resistance of $R_3$ in $k\Omega$ because according to Ohm's Law $R = \frac{V}{I}$.

\section*{Conclusion}

The linear relationship between voltage and current through an ohmic resistor is an important concept in circuits. 

\end{document}